Multilevel computation of glacier geometry \\ from Stokes dynamics

### 2021-06-07 paper/msg.tex
Nothing fiddles.

@@

FIXME MCD = multilevel constraint decomposition, a multigrid \cite{Trottenbergetal2001} method basically by \cite{Tai2003} for obstacle problems; see \cite{Bueler2022};  obstacle problem view first extended to Stokes by \cite{WirbelJarosch2020};  The computations in this paper use the Python FE library Firedrake \cite{Rathgeberetal2016}, which applies an embedded domain language \cite{Alnaesetal2014} to convert weak forms into discrete equations which are solved in parallel using the PETSc \cite{Balayetal2020} library.

@@

\begin{figure}[t]
\begin{center}
FIXME %\includegraphics[width=\textwidth]{genfigs/extruded.pdf}
\end{center}
\caption{A sketch of the mass-perturbation approximation in equations \eqref{eq:systemmassanalogy} and \eqref{eq:massanalogy}.}
\label{fig:massanalogy}
\end{figure}

@@

\begin{pcode}[ht]
\begin{pseudo*}
\pr{mcdn-solver}(J,s^J,b^J,\id{atol}=10^{-20},\id{rtol}=10^{-3},\id{cyclemax}=100)\text{:} \\+
    $\rho_0=\vertiii{(a,\cdot) - F^J(s^J)[\cdot]}_{(s^J,b^J)}$ \\
    for $k=1,\dots,\id{cyclemax}$ \\+
        $s^J \gets s^J + \pr{mcdn-vcycle}(J,s^J,b^J)$ \\
        $\rho_k=\vertiii{(a,\cdot) - F^J(s^J)[\cdot]}_{(s^J,b^J)}$ \\
        if $\rho_k < \id{atol}$ or $\rho_k \le \id{rtol} \, \rho_0$ \\+
            break \\--
\end{pseudo*}
\caption{The SIGP is solved by iterating V-cycles (Pseudocode \ref{pc:mcdn-vcycle}) until the CP residual norm \eqref{eq:cpnorm} is small.}
\label{pc:mcdn-solver}
\end{pcode}

@@

FIXME in this paper the Glen-Stokes problem is solved by Newton linearization and (parallel) direct solution of the Newton step equations; Firedrake and PETSc \cite{Balayetal2020,Bueler2021} used to solve this dynamics problem and thereby evaluate the residual $F$; geometric multigrid schemes exist for the  Glen-Stokes problem \cite{IsaacStadlerGhattas2015}; see also \cite{BrownSmithAhmadia2013} and \cite{Tuminaroetal2016} which solve a first-order approximation of the Glen-Stokes ice sheet problem by geometric and algebraic multigrid, respectively

@@

\section*{Acknowledgments}  Thanks to David Maxwell for discussions regarding the formulation and well-posedness of the model.

>>>>

\date{\today}

@@

FIXME a multigrid \cite{Trottenbergetal2001} method basically by \cite{Tai2003} for obstacle problems; see \cite{Bueler2022} and \cite{GraeserKornhuber2009};  obstacle problem view first extended to Stokes by \cite{WirbelJarosch2020}

@@

\begin{figure}[t]
\begin{center}
FIXME %\includegraphics[width=\textwidth]{genfigs/massanalogy.pdf}
\end{center}
\caption{A sketch of the mass-perturbation approximation in equations \eqref{eq:systemmassanalogy} and \eqref{eq:massanalogy}.}
\label{fig:massanalogy}
\end{figure}

@@

\begin{pcode}[ht]
\begin{pseudo*}
\pr{mcdn-solver}(J,s^J,b^J,\id{atol}=10^{-20},\id{rtol}=10^{-3},\id{stol}=10^{-20},\id{max}=100)\text{:} \\+
    $\rho_0=\vertiii{(a,\cdot) - F^J(s^J)[\cdot]}_{(s^J,b^J)}$ \\
    for $k=1,\dots,\id{cyclemax}$ \\+
        $z^J = \pr{mcdn-vcycle}(J,s^J,b^J)$ \\
        $s^J \gets s^J + z^J$ \\
        $\rho_k=\vertiii{(a,\cdot) - F^J(s^J)[\cdot]}_{(s^J,b^J)}$ \\
        if $\rho_k < \id{atol}$ or $\rho_k \le \id{rtol} \, \rho_0$ or $\|z^J\| \le \id{stol}$ \\+
            break \\--
\end{pseudo*}
\caption{The SIGP is solved in-place by iterating V-cycles (Pseudocode \ref{pc:mcdn-vcycle}) until the CP residual norm \eqref{eq:cpnorm} is small.}
\label{pc:mcdn-solver}
\end{pcode}

@@

FIXME The computations in this section use the Python FE library Firedrake \cite{Rathgeberetal2016}, which applies an embedded domain language \cite{Alnaesetal2014} to convert weak forms into discrete equations which are solved in parallel using the PETSc \cite{Balayetal2020} library.  Each time the smoother is applied the Glen-Stokes problem is solved, by Newton linearization and (parallel) direct solution of the Newton step equations, in order to evaluate the IDO $\Phi(s)$.  To replace this non-scalable approach we propose in future work to apply existing geometric \cite{BrownSmithAhmadia2013,IsaacStadlerGhattas2015} and algebraic \cite{Tuminaroetal2016} multigrid strategies.

@@

\section*{Acknowledgments}  Thanks to David Maxwell for discussions regarding the formulation of the model.
